---
title: "Analysis of Gryllus bimaculatus density Oct 23 data"
author: "Edouard Bessette"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_float: yes
  word_document:
    toc: yes
---

```{r setup, include=FALSE}
library(formatR)
library(knitr)
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, tidy = TRUE, tidy.opts=list(width.cutoff=60),  fig.width = 6, fig.asp = 0.8, out.width = "80%")
```

# Load the necessary libraries

```{r }
library(readxl)
library(dplyr)
library(ggplot2)
library(survminer)
library(survival)
library(RVAideMemoire)
library(lmtest)
library(lme4)
library(flexsurv)
library(car)
library(emmeans)
library(multcomp)
library(Hmisc)
library(extrafont)
loadfonts()
setwd("C:/Users/eb826/OneDrive - University of Exeter/PhD_EdouardMicrosporidia/Thesis/Chapter5_Gryllus bimaculatus/Group Experiments")
```

# Survival analysis

## Import data

```{r }
survival_data_Gbimaculatus <- read_excel("GbdensityOct23_survival_data.xlsx")
colnames(survival_data_Gbimaculatus)
```
## Convert factors to appropriate levels

```{r }
survival_data_Gbimaculatus[c("Treatment", "Infection", "Density", "R", "Box_ID")] <- lapply(survival_data_Gbimaculatus[c("Treatment", "Infection", "Density", "R", "Box_ID")], factor)
```

## Reorganise levels of Treatment variable and convert variables to factors

```{r }
survival_data_Gbimaculatus$Treatment <- factor(survival_data_Gbimaculatus$Treatment,levels = c("N_8","Y_8","N_16","Y_16","N_32","Y_32"))
```

## Data summary

```{r }
str(survival_data_Gbimaculatus)
summary(survival_data_Gbimaculatus)
```

## Check mortality (sanity check)

```{r }
# Calculate total number of crickets and number of deaths by treatment
total_Gbimaculatus <- survival_data_Gbimaculatus %>% 
  group_by(Treatment) %>% 
  summarise(total = n())

mortality_Gbimaculatus <- survival_data_Gbimaculatus %>% 
  group_by(Treatment, Censor) %>% 
  summarise(mortality = sum(Censor == 1)) %>%
  filter(Censor == 1) # Keep only dead crickets

# Merge total and mortality data frames
Gbimaculatus_check <- merge(total_Gbimaculatus, mortality_Gbimaculatus, by = "Treatment", all = TRUE)

# Calculate percentage of mortality
Gbimaculatus_check$percent <- (Gbimaculatus_check$mortality / Gbimaculatus_check$total) * 100
print(Gbimaculatus_check)

# Create a data frame with the required columns
contingency_data <- data.frame(Treatment = Gbimaculatus_check$Treatment,
                               Alive = Gbimaculatus_check$total - Gbimaculatus_check$mortality,
                               Dead = Gbimaculatus_check$mortality)

# Set Treatment as row names
row.names(contingency_data) <- contingency_data$Treatment
contingency_data$Treatment <- NULL

# Reorder rows according to desired order
desired_order <- c("N_8", "Y_8", "N_16", "Y_16", "N_32", "Y_32")
contingency_data <- contingency_data[match(desired_order, rownames(contingency_data)), ]
print(contingency_data)

# Perform chi-square test
chi_square_test <- chisq.test(contingency_data)
print(chi_square_test)

# Perform pairwise comparisons using fdr correction
contingency_matrix <- as.matrix(contingency_data)
pairwise_chisq <- pairwise.prop.test(x = contingency_matrix, p.adjust.method = "fdr")
print(pairwise_chisq)
```

## Surv curves by treatments

```{r }
surv <- survfit(Surv(Death_time, Censor) ~ Treatment, data = survival_data_Gbimaculatus)
surv_plot <- ggsurvplot(
  fit = surv,
  surv.scale = "percent",
  title = "Cricket survival per treatment",
  legend.title = "Exposure_Density",
  legend.labs = c("N_8","Y_8","N_16","Y_16","N_32","Y_32"),
  palette = c("#c4c4ff","#ffc4c4","#0000ff","#ff4e4e","#000027","#760000"),
  linetype = c(1, 2, 1, 2, 1, 2),
  ggtheme = theme_bw(base_family = "Times New Roman") +
    theme(
      axis.text = element_text(size = 12), 
      axis.title = element_text(size = 12, vjust = 0.5),
      legend.text = element_text(size = 10),
      legend.title = element_text(size = 10, face = "bold"),
      legend.direction = "horizontal",
      plot.title = element_text(size = 12, face = "bold", hjust = 0.5)))
surv_plot

# Get p-values with coordinates
#pvalue_letters <- surv_pvalue(surv, method = "survdiff", get_coord = TRUE)
```

## LT50 determined with median values

```{r }
summary(surv)$table
```

## Determine if we have a constant risk or not

```{r }
plotsurvivors(survival_data_Gbimaculatus$Death_time, survival_data_Gbimaculatus$Censor)
# The instantaneous risk is considered constant if the points on the survival curve are roughly aligned on a straight line.
## Here the instant risk is constant, then we opt for a survival regression model
```

## Surv reg model

### Find the best distribution for our survival data
```{r}
# Create vector of parametric distributions to test
distList <- c(distList <- c("extreme", "logistic", "gaussian", "weibull", "exponential", "rayleigh", "lognormal", "loglogistic", "t"))

# Function fits each distribution and extracts AIC, BIC, log-likelihood values
## The distribution is estimated on the event data without any covariate (~ 1)
fit_dist <- function(dist) {
  tmp <- survreg(Surv(Death_time, Censor) ~ 1, data = survival_data_Gbimaculatus, dist = dist)
  c(AIC(tmp), BIC(tmp), as.numeric(logLik(tmp)))
}

# Apply above function to each distribution in the distList
results_list <- lapply(distList, fit_dist)

# Convert the above list of results to a data frame
results_df <- data.frame(t(matrix(unlist(results_list), nrow = 3)))
colnames(results_df) <- c("aic", "bic", "logLik")
rownames(results_df) <- distList

# Find the distribution with the lowest AIC, BIC, and logLik values
bestFitAIC <- rownames(results_df)[which.min(results_df$aic)]
bestFitBIC <- rownames(results_df)[which.min(results_df$bic)]
bestFitLogLik <- rownames(results_df)[which.max(results_df$logLik)]

# Print the results data frame and best fitting distributions
results_df
cat("\nBest fitting distribution using AIC:", bestFitAIC, "\n")
cat("\nBest fitting distribution using BIC:", bestFitBIC, "\n")
cat("\nBest fitting distribution using Log-Likelihood:", bestFitLogLik, "\n")
## The best distribution here is lognormal

# We can check graphically how the distribution is fitting to the survival data
require(flexsurv)
s <- with(survival_data_Gbimaculatus,Surv(Death_time,Censor))
sLnorm  <- flexsurvreg(s ~ 1,dist="lnorm",data=survival_data_Gbimaculatus)
sExp  <- flexsurvreg(s ~ 1,dist="exponential",data=survival_data_Gbimaculatus) 
plot(sLnorm) +
lines(sExp, col="blue")
```

```{r }
model <- survreg(Surv(Death_time, Censor) ~ Infection * Density, data = survival_data_Gbimaculatus, dist="lognormal") # distribution determined earlier
RVAideMemoire::plotresid(model)
summary(model)
Anova(model, type="II")
# According to the model ANOVA, density had a significant impact on survival
```

## Pairwise differences between different treatments

```{r include=FALSE}
pairwise_comp_i <- emmeans(model, pairwise ~ Infection, type="response")
pairwise_comp_i
cld(pairwise_comp_i,adjust="fdr",Letters=letters)

pairwise_comp <- emmeans(model, pairwise ~ Density, type="response")
pairwise_comp
cld(pairwise_comp,adjust="fdr",Letters=letters)

# Even if the ANOVA showed that the infection had no effect, some comparison within the density treatment could reveal some infection effect
pairwise_comp1 <- emmeans(model, pairwise ~ Density:Infection, type="response")
pairwise_comp1
cld(pairwise_comp1,adjust="fdr",Letters=letters)
```

# Emergence Analysis

## Import data for both sexes

### For this analysis, I made three datasets. Emergence_Time represents the day of emergence during the cricket life and Censor the event of emergence (0,1). One dataset has a Sex variable with 'Male', 'Female' and 'NA' for the individuals that did not emerge (Censor = 0). Then I made one dataset only with the Males individual (Females and non-emerged recorded as Censor = 0) and one dataset for the females.

```{r }
emergence_data_Gbimaculatus <- read_excel("AdultEmergence_data.xlsx")
colnames(emergence_data_Gbimaculatus)
```

### Convert factors to appropriate levels

```{r }
emergence_data_Gbimaculatus[c("Treatment","Infection","Density","R","Sex")] <- lapply(emergence_data_Gbimaculatus[c("Treatment","Infection","Density","R","Sex")],factor)
```

### Reorganise levels of Treatment variable and convert variables to factors

```{r }
emergence_data_Gbimaculatus$Treatment <- factor(emergence_data_Gbimaculatus$Treatment,levels = c("N_8","Y_8","N_16","Y_16","N_32","Y_32"))
```

### Data summary

```{r }
str(emergence_data_Gbimaculatus)
```

## Import Male dataset

```{r }
emergence_male_Gbimaculatus <- read_excel("MaleEmergence_data.xlsx")
colnames(emergence_male_Gbimaculatus)
```

### Convert factors to appropriate levels

```{r }
emergence_male_Gbimaculatus[c("Treatment","Infection","Density","R")] <- lapply(emergence_male_Gbimaculatus[c("Treatment","Infection","Density","R")],factor)
```

### Reorganise levels of Treatment variable and convert variables to factors

```{r }
emergence_male_Gbimaculatus$Treatment <- factor(emergence_male_Gbimaculatus$Treatment,levels = c("N_8","Y_8","N_16","Y_16","N_32","Y_32"))
```

### Data summary

```{r }
str(emergence_male_Gbimaculatus)
```

## Import Female dataset

```{r }
emergence_female_Gbimaculatus <- read_excel("FemaleEmergence_data.xlsx")
colnames(emergence_female_Gbimaculatus)
```

### Convert factors to appropriate levels

```{r }
emergence_female_Gbimaculatus[c("Treatment","Infection","Density","R")] <- lapply(emergence_female_Gbimaculatus[c("Treatment","Infection","Density","R")],factor)
```

### Reorganise levels of Treatment variable and convert variables to factors

```{r }
emergence_female_Gbimaculatus$Treatment <- factor(emergence_female_Gbimaculatus$Treatment,levels = c("N_8","Y_8","N_16","Y_16","N_32","Y_32"))
```

### Data summary

```{r }
str(emergence_female_Gbimaculatus)
```

## Calculate proportions and perform chi-square tests

```{r echo=TRUE, message=FALSE}
emergence_table <- emergence_data_Gbimaculatus %>%
  group_by(Treatment,Infection,Density) %>%
  summarise(y1 = sum(Censor == 1), #Number of emerged crickets
            y2 = sum(Censor == 0)   #Number of non-emerged crickets
  ) %>%
  mutate(y1_prop = y1 / (y1 + y2), #Proportion of emerged crickets
         y2_prop = y2 / (y1 + y2))   #Proportion of non-emerged crickets

print(emergence_table)

# Create the contingency table
contingency_table <- as.matrix(emergence_table[, c("y1", "y2")])
rownames(contingency_table) <- paste(emergence_table$Treatment)
colnames(contingency_table) <- c("Emerged", "Non-emerged")

# Perform chi-square test
chisq_result <- chisq.test(contingency_table)
print(chisq_result)

# Perform pairwise comparisons using fdr correction
pairwise_chisq <- pairwise.prop.test(x = contingency_table, p.adjust.method = "fdr")
print(pairwise_chisq)
```

## LT50 table

```{r }
surv_emergence <- survfit(Surv(Emergence_time,Censor) ~ Sex,data = emergence_data_Gbimaculatus)
summary(surv_emergence)$table
```

## Surv curve emergence

```{r }
surv_emergence <- survfit(Surv(Emergence_time,Censor) ~ Treatment,data = emergence_data_Gbimaculatus)
surv_plot_emergence <- ggsurvplot(
  fit = surv_emergence,
  surv.scale = "percent",
  title = "Cricket emergence",
  ylab = "Cricket emergence",
  legend.title = "Exposure_Density",
  legend.labs = c("N_8","Y_8","N_16","Y_16","N_32","Y_32"),
  palette = c("#c4c4ff","#ffc4c4","#0000ff","#ff4e4e","#000027","#760000"),
  linetype = c(1, 2, 1, 2, 1, 2),
  xlim = c(28,50),
  break.x.by = 4,
  fun="event",
  #surv.median.line = "h",
  ggtheme = theme_bw(base_family = "Times New Roman") +
    theme(
      axis.text = element_text(size = 12), 
      axis.title = element_text(size = 12, vjust = 0.5),
      legend.text = element_text(size = 10),
      legend.title = element_text(size = 10, face = "bold"),
      plot.title = element_text(size = 12, face = "bold", hjust = 0.5)))

# Add line at 50% emergence
p1 <- surv_plot_emergence$plot +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "gold")
p1
```

## Surv curve male

```{r }
surv_male <- survfit(Surv(Emergence_time,Censor) ~ Treatment,data = emergence_male_Gbimaculatus)
surv_plot_male <- ggsurvplot(
  fit = surv_male,
  surv.scale = "percent",
  title = "Cricket male emergence",
  ylab = "Cricket emergence",
  legend = "none",
  #legend.title = "Exposure_Density",
  #legend.labs = c("N_8","Y_8","N_16","Y_16","N_32","Y_32"),
  palette = c("#c4c4ff","#ffc4c4","#0000ff","#ff4e4e","#000027","#760000"),
  linetype = c(1, 2, 1, 2, 1, 2),
  xlim = c(28,50),
  fun="event",
  ggtheme = theme_bw(base_family = "Times New Roman") +
    theme(
      axis.text = element_text(size = 12), 
      axis.title = element_text(size = 12, vjust = 0.5),
      legend.text = element_text(size = 10),
      legend.title = element_text(size = 10),
      plot.title = element_text(size = 12, face = "bold", hjust = 0.5)))
surv_plot_male
```

## Surv curve female

```{r }
surv_female <- survfit(Surv(Emergence_time,Censor) ~ Treatment,data = emergence_female_Gbimaculatus)
surv_plot_female <- ggsurvplot(
  fit = surv_female,
  surv.scale = "percent",
  title = "Cricket female emergence",
  ylab = "Cricket emergence",
  legend = "none",
  #legend.title = "Exposure_Density",
  #legend.labs = c("N_8","Y_8","N_16","Y_16","N_32","Y_32"),
  palette = c("#c4c4ff","#ffc4c4","#0000ff","#ff4e4e","#000027","#760000"),
  linetype = c(1, 2, 1, 2, 1, 2),
  xlim = c(28,50),
  ylim = c(0,0.4),
  fun="event",
  ggtheme = theme_bw(base_family = "Times New Roman") +
    theme(
      axis.text = element_text(size = 12), 
      axis.title = element_text(size = 12, vjust = 0.5),
      legend.text = element_text(size = 10),
      legend.title = element_text(size = 10),
      plot.title = element_text(size = 12, face = "bold", hjust = 0.5)))
surv_plot_female
```

## Arrange the surv plots together

```{r }
splots <- list(surv_plot_female, surv_plot_male)
arrange <- arrange_ggsurvplots(splots, print = TRUE, ncol = 2, nrow = 1)
arrange
```

## Generalized linear model – binomial family

```{r }
#Filter out NA values from the initial data set
emergence_data_filtered <- emergence_data_Gbimaculatus %>% filter(Sex != "NA")

#Compare models with and without Sex
glm_model <- glm(Emergence_time ~ Infection * Density + Sex, data = (emergence_data_filtered),family = Gamma(link = "inverse"))
glm_model1 <- glm(Emergence_time ~ Infection * Density, data = (emergence_data_filtered),family = Gamma(link = "inverse"))
anova(glm_model, glm_model1)
#We keep model without sex
RVAideMemoire::plotresid(glm_model1)

summary(glm_model1)
Anova(glm_model1,type="II")

pairwise_comp2 <- emmeans(glm_model1,pairwise ~ Infection,type="response")
pairwise_comp2
cld(pairwise_comp2,adjust="fdr",Letters=letters)
```

## Here is a geombar representation of the emergence if the surv curves are not adapted

### Calculate mean and standard deviation

```{r }
summary_data <- emergence_data_filtered %>%
  group_by(Treatment, Sex) %>%
  summarise(mean_time = mean(Emergence_time),
            sd_time = sd(Emergence_time))
```

### Define lighter colors

```{r }
light_blue <- alpha("blue", 0.5)
light_red <- alpha("red", 0.5)
```

### Plot Emergence_time vs. Treatment colored by Sex with bars and mean/SD summary

```{r }
plot <- ggplot(data = emergence_data_filtered, aes(x = Treatment, y = Emergence_time, fill = Sex)) +
  geom_bar(stat = "summary", fun = "mean", position = "dodge", alpha = 0.5, color = "black", show.legend = FALSE) +
  stat_summary(fun.data = mean_sdl, geom = "point", position = position_dodge(width = 0.75), size = 3, color = "black", show.legend = FALSE) +
  geom_errorbar(data = summary_data, aes(ymin = mean_time - sd_time, ymax = mean_time + sd_time, y = mean_time), #Error bars for SD
                position = position_dodge(width = 0.75), width = 0.2) +
  labs(x = "Treatment", y = "Emergence Time", title = "Mean and SD of Emergence Time by Treatment and Sex") + 
  facet_grid(~ Sex) + #Facet by Sex
  scale_fill_manual(values = c(light_blue, light_red)) +
  theme_bw(base_family = "Times New Roman") + 
  theme(axis.text = element_text(size = 12), 
        axis.title = element_text(size = 12), 
        plot.title = element_text(size = 12, face = "bold", hjust = 0.5),
        strip.text = element_text(size = 12))  #Adjust facet titles font size
plot
```

# Weight Analysis

## Import data

```{r }
weight_data_Gbimaculatus <- read_excel("Weight_data.xlsx")
colnames(weight_data_Gbimaculatus)
```

## Reorganise levels of Treatment variable and convert variables to factors

```{r }
weight_data_Gbimaculatus[c("Treatment", "R", "Sex", "Infection", "Density", "Ind")] <- lapply(weight_data_Gbimaculatus[c("Treatment", "R", "Sex", "Infection", "Density", "Ind")], factor)
weight_data_Gbimaculatus$Treatment <- factor(weight_data_Gbimaculatus$Treatment,
                                             levels = c("N_8", "Y_8", 
                                                        "N_16", "Y_16", "N_32",
                                                        "Y_32"))
str(weight_data_Gbimaculatus)
```

## Calculate mean and standard deviation for weight variable by treatment

```{r echo=TRUE}
summary_data_w <- aggregate(Weight ~ Treatment * Sex, weight_data_Gbimaculatus, FUN = function(x) c(mean = mean(x), sd = sd(x)))
print(summary_data_w)
```

## Box plot of Weight by Treatment

```{r }
boxplot_weight <- ggplot(weight_data_Gbimaculatus, aes(x = Treatment, y = Weight, color = Treatment),
  legend.title = "Exposure_Density",
  legend.labs = c("N_8","Y_8","N_16","Y_16","N_32","Y_32")) +
  geom_boxplot() +
  geom_point() +
  scale_color_manual(values = c("#c4c4ff","#ffc4c4","#0000ff","#ff4e4e","#000027", "#760000")) +
  labs(title = "Maturation weight by treatment",
       x = "Treatment",
       y = "Weight (mg)") +
  theme_bw(base_family = "Times New Roman") + 
  theme(axis.text = element_text(size = 12), 
        axis.title = element_text(size = 12), 
        axis.text.x = element_text(angle = 50, vjust = 0.6),
        legend.title = element_text(size = 10, face = "bold"),
        plot.title = element_text(size = 12, face = "bold", hjust = 0.5))
```

## Box Plot of weight Facet by Sex

```{r }
boxplot_faceted <- boxplot_weight + facet_wrap(~ Sex) +
  theme(strip.text.x = element_text(size=12)) +
  labs(colour = "Exposure_Density")
boxplot_faceted
```

## General model
```{r }
hist(weight_data_Gbimaculatus$Weight)

model_weight <- lm(Weight ~ Infection * Density + Sex, data = weight_data_Gbimaculatus)
model_weight1 <- lm(Weight ~ Infection * Density * Sex, data = weight_data_Gbimaculatus)

anova(model_weight,model_weight1)
```

### Check residual and test significance of each term of the model_weight with Anova

```{r }
RVAideMemoire::plotresid(model_weight1)
summary(model_weight1)

Anova(model_weight1, type="II")
```

### Pairwise comparisons

```{r }
pairwise_comp_w <- emmeans(model_weight1, pairwise ~ Infection, type="response")
pairwise_comp_w
cld(pairwise_comp_w, adjust="fdr", Letters=letters)

pairwise_comp_w1 <- emmeans(model_weight1, pairwise ~ Density, type="response")
pairwise_comp_w1
cld(pairwise_comp_w1, adjust="fdr", Letters=letters)

pairwise_comp_w2 <- emmeans(model_weight1, pairwise ~ Sex, type="response")
pairwise_comp_w2
cld(pairwise_comp_w2, adjust="fdr", Letters=letters)

pairwise_comp_w3 <- emmeans(model_weight1, pairwise ~ Sex*Infection, type="response")
pairwise_comp_w3
cld(pairwise_comp_w3, adjust="fdr", Letters=letters)
```

# Feed and feces analysis

## Import data

```{r }
feed_feces_data_Gbimaculatus <- read_excel("Feed_Feces_data.xlsx")
colnames(feed_feces_data_Gbimaculatus)
summary(feed_feces_data_Gbimaculatus)
```

## Reorganise levels of Treatment variable and convert variables to factors
### Here the experimental box was added as random factor, as multiple measurments (paired serie) were recorded on it

```{r }
feed_feces_data_Gbimaculatus[c("Treatment", "R", "Box_Id", "Infection", "Density", "Dpi", "Cricket_Age")] <- lapply(feed_feces_data_Gbimaculatus[c("Treatment", "R", "Box_Id", "Infection", "Density", "Dpi", "Cricket_Age")], factor)
feed_feces_data_Gbimaculatus$Treatment <- factor(feed_feces_data_Gbimaculatus$Treatment,
                                                 levels = c("N_8","Y_8","N_16","Y_16","N_32","Y_32"))
str(feed_feces_data_Gbimaculatus)
```

## Calculate mean and standard deviation for each response variable by treatment

```{r }
summary_data_f <- aggregate(cbind(Feces_ind_day, Feed_ind_day) ~ Treatment, feed_feces_data_Gbimaculatus, FUN = function(x) c(mean = mean(x), sd = sd(x)))
print(summary_data_f)
```

## Boxplot of Feed Consumed by Treatment

```{r }
boxplot_feed <- ggplot(feed_feces_data_Gbimaculatus, aes(x = Treatment, y = Feed_ind_day, color = Treatment),
  legend.title = "Exposure_Density",
  legend.labs = c("N_8","Y_8","N_16","Y_16","N_32","Y_32")) +
  geom_boxplot() +
  geom_point() +
  scale_color_manual(values = c("#c4c4ff","#ffc4c4","#0000ff","#ff4e4e","#000027", "#760000")) +
  labs(title = "Feed consumed by treatment",
       x = "Treatment",
       y = "Feed consumed per individual and per day (mg)") +
  theme_bw(base_family = "Times New Roman") + 
  theme(axis.text = element_text(size = 12), 
        axis.title = element_text(size = 12), 
        plot.title = element_text(size = 12, face = "bold", hjust = 0.5))
boxplot_feed
```

## Boxplot of Feces Produced by Treatment

```{r }
boxplot_feces <- ggplot(feed_feces_data_Gbimaculatus, aes(x = Treatment, y = Feces_ind_day, color = Treatment),
  legend.title = "Exposure_Density",
  legend.labs = c("N_8","Y_8","N_16","Y_16","N_32","Y_32")) +
  geom_boxplot() +
  geom_point() +
  scale_color_manual(values = c("#c4c4ff","#ffc4c4","#0000ff","#ff4e4e","#000027", "#760000")) +
  labs(title = "Faeces produced by treatment",
       x = "Treatment",
       y = "Faeces per individual and per day (mg)") +
  theme_bw(base_family = "Times New Roman") + 
  theme(axis.text = element_text(size = 12), 
        axis.title = element_text(size = 12), 
        plot.title = element_text(size = 12, face = "bold", hjust = 0.5))
boxplot_feces
```

## Boxplot of Feces by Treatment and Dpi (obsolete)

```{r }
# Calculate the log transformation of Feces_ind_day
#feed_feces_data_Gbimaculatus$log_Feces_ind_day <- log(feed_feces_data_Gbimaculatus$Feces_ind_day)

# Create boxplot with log-transformed y-axis
boxplot_feces <- ggplot(feed_feces_data_Gbimaculatus, aes(x = Treatment, y = Feces_ind_day, color = Treatment)) +
  geom_boxplot(outlier.shape = NA, size = 1) +
  scale_color_manual(values = c("#c4c4ff","#ffc4c4","#0000ff","#ff4e4e","#000027", "#760000")) +
  labs(title = "Faeces produced by Treatment and Dpi",
       x = "Treatment",
       y = "Faeces per individual and per day (mg)") +
  facet_wrap(~Dpi) +  
  theme_bw(base_family = "Times New Roman") + 
  theme(axis.text = element_text(size = 12), 
        axis.title = element_text(size = 12),
        axis.text.x = element_text(angle = 50, vjust = 0.6),
        legend.title = element_text(size = 10, face = "bold"),
        plot.title = element_text(size = 12, face = "bold", hjust = 0.5)) +
  labs(colour = "Exposure_Density")

boxplot_feces
```

## Correlation between consumed feed and Faeces

```{r }
correlation_plot <- ggplot(feed_feces_data_Gbimaculatus, aes(x = Feces_ind_day, y = Feed_ind_day)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Correlation between consumed feed and faeces",
       x = "Faeces per individual and per day (mg)",
       y = "Feed consumed per individual and per day (mg)") +
  theme_bw(base_family = "Times New Roman") + 
 theme(plot.title = element_text(face = "bold", size = 12, hjust = 0.5),
        axis.title.y = element_text(size = 12), 
        axis.title.x = element_text(size = 12),  
        axis.text.x = element_text(angle = 50, vjust = 1, hjust = 1, size = 12),
        axis.text.y = element_text(size = 12),  
        legend.position = "none")

correlation_plot
```

### Compute correlation coefficient

```{r }
correlation <- cor(feed_feces_data_Gbimaculatus$Feed_ind_day, feed_feces_data_Gbimaculatus$Feces_ind_day)
```

### Conduct Pearson's correlation test

```{r }
cor_test <- cor.test(feed_feces_data_Gbimaculatus$Feed_ind_day, feed_feces_data_Gbimaculatus$Feces_ind_day)
```

### Print correlation coefficient

```{r }
print(paste("Correlation coefficient:", correlation))
```

### Print correlation test results

```{r }
print(cor_test)
```

## Linear Model (lm) for biomass and Quantitative continuous variables

### Feces model

```{r }
hist(feed_feces_data_Gbimaculatus$Feces_ind_day)

model_feces <- lmer(Feces_ind_day ~ Infection * Density + (1|Box_Id), data = feed_feces_data_Gbimaculatus) #The box id is a random effect (different box measured across different time points)

model_feces_age <- lmer(Feces_ind_day ~ Infection * Density * Cricket_Age + (1|Box_Id), data = feed_feces_data_Gbimaculatus)

anova(model_feces, model_feces_age)
# The model including ages has lower AIC and BIC
```

#### Check residuals and compare the models and significance of each term of the model with Anova

```{r }
RVAideMemoire::plotresid(model_feces_age)
summary(model_feces_age)
Anova(model_feces_age, type="II")
```

#### Pairwise comparisons

```{r }
pairwise_comp_faeces <- emmeans(model_feces_age, pairwise ~ Infection, type="response")
pairwise_comp_faeces
cld(pairwise_comp_faeces, adjust="fdr", Letters=letters)

pairwise_comp_faeces1 <- emmeans(model_feces_age, pairwise ~ Cricket_Age, type="response")
pairwise_comp_faeces1
cld(pairwise_comp_faeces1, adjust="fdr", Letters=letters)

pairwise_comp_faeces2 <- emmeans(model_feces_age, pairwise ~ Infection*Cricket_Age, type="response")
pairwise_comp_faeces2
cld(pairwise_comp_faeces2, adjust="fdr", Letters=letters)
```

### Feed model

```{r }
model_feed <- lmer(Feed_ind_day ~ Infection * Density + (1|Box_Id), data = feed_feces_data_Gbimaculatus) #The box id is a random effect (different box measured across different time points
model_feed_dpi <- lmer(Feed_ind_day ~ Infection * Density + Dpi + (1|Box_Id), data = feed_feces_data_Gbimaculatus)
```

#### Check residuals

```{r }
RVAideMemoire::plotresid(model_feed_dpi)
summary(model_feed_dpi)
```

#### Test significance of each term of the model with Anova

```{r }
Anova(model_feed_dpi, type="II")
```

#### Pairwise comparison

```{r }
pairwise_comp_feed <- emmeans(model_feed_dpi, pairwise ~ Infection|Dpi, type="response")
pairwise_comp_feed
cld(pairwise_comp_feed, adjust="fdr", Letters=letters)
```

# Fumagillin assay

## Import data

```{r }
survival_data_Fumagillin <- read_excel("Fumagillin_survival_data.xlsx")
colnames(survival_data_Fumagillin)
summary(survival_data_Fumagillin)
```

## Convert factors to appropriate levels

```{r }
survival_data_Fumagillin[c("Treatment", "R")] <- lapply(survival_data_Fumagillin[c("Treatment", "R")], factor)
str(survival_data_Fumagillin)
```

## Check mortality

```{r }
# Calculate total number of crickets and number of deaths by treatment
total_Fumagillin <- survival_data_Fumagillin %>% 
  group_by(Treatment) %>% 
  summarise(total = n())

mortality_Fumagillin <- survival_data_Fumagillin %>% 
  group_by(Treatment, Censor) %>% 
  summarise(mortality = sum(Censor == 1)) %>%
  filter(Censor == 1) # Keep only dead crickets

# Merge total and mortality data frames
Fumagillin_check <- merge(total_Fumagillin, mortality_Fumagillin, by = "Treatment", all = TRUE)

# Calculate percentage of mortality
Fumagillin_check$percent <- (Fumagillin_check$mortality / Fumagillin_check$total) * 100
print(Fumagillin_check)

# Create a data frame with the required columns
contingency_fuma <- data.frame(Treatment = Fumagillin_check$Treatment,
                               Alive = Fumagillin_check$total - Fumagillin_check$mortality,
                               Dead = Fumagillin_check$mortality)

# Set Treatment as row names
row.names(contingency_fuma) <- contingency_fuma$Treatment
contingency_fuma$Treatment <- NULL

# Perform chi-square test
chisq_result <- chisq.test(contingency_fuma)
print(chisq_result)
```

## Surv curve Treatment

```{r }
surv_fumagillin <- survfit(Surv(Death_time, Censor) ~ Treatment, data = survival_data_Fumagillin)
surv_plot_fuma <- ggsurvplot(fit = surv_fumagillin,
  conf.int = TRUE,
  #conf.int.style = "step",
  conf.int.alpha = 0.1, 
  surv.scale = "percent",
  surv.median.line = "hv",
  title = "Cricket survival per treatment (with CI)",
  legend.title = "Treatments",
  legend.labs = c("Fumagillin","Untreated"),
  palette = c("#c4c4ff","#ffc4c4"),
  ggtheme = theme_bw(base_family = "Times New Roman") +
    theme(axis.text = element_text(size = 12), 
      axis.title = element_text(size = 12, vjust = 0.5),
      legend.text = element_text(size = 10),
      legend.title = element_text(size = 10, face = "bold"),
      legend.direction = "horizontal",
      plot.title = element_text(size = 12, face = "bold", hjust = 0.5)))

# Get p-values as letters with coordinates
pvalue_letters <- surv_pvalue(surv_fumagillin, get_coord = TRUE)

# Add p-values as annotations
surv_plot_fuma$plot <- surv_plot_fuma$plot +
  ggplot2::annotate("text", x = pvalue_letters$pval.x, y = pvalue_letters$pval.y, label = pvalue_letters$pval.txt, size = 4, color = "black")

# Display the plot with p-values
surv_plot_fuma$plot
```

## LT50 table

```{r }
summary(surv_fumagillin)$table
```

## Surv model

### Determine if we have a constant risk or not

```{r }
RVAideMemoire::plotsurvivors(survival_data_Fumagillin$Death_time, survival_data_Fumagillin$Censor)
# The instantaneous risk is considered constant if the points on the survival curve are roughly aligned on a straight line. Here, the risk seems constant, but with a drop at the end. Survreg model seems the more adapted.
```

### Find the best distribution for our survival data
```{r}
# Create vector of parametric distributions to test
## To adapt depending on the model you use (survreg, cox, flexsurvreg...) as the dist names can change
distList <- c(distList <- c("extreme", "logistic", "gaussian", "weibull", "exponential", "rayleigh", "lognormal", "loglogistic", "t"))

# Function fits each distribution and extracts AIC, BIC, log-likelihood values
## The distribution is estimated on the event data without any covariate (~ 1) / intercept only
fit_dist <- function(dist) {
  tmp <- survreg(Surv(Death_time, Censor) ~ 1, data = survival_data_Fumagillin, dist = dist)
  c(AIC(tmp), BIC(tmp), as.numeric(logLik(tmp)))
}

# Apply above function to each distribution in the distList
results_list <- lapply(distList, fit_dist)

# Convert the above list of results to a data frame
results_df <- data.frame(t(matrix(unlist(results_list), nrow = 3)))
colnames(results_df) <- c("aic", "bic", "logLik")
rownames(results_df) <- distList

# Find the distribution with the lowest AIC, BIC, and logLik values
bestFitAIC <- rownames(results_df)[which.min(results_df$aic)]
bestFitBIC <- rownames(results_df)[which.min(results_df$bic)]
bestFitLogLik <- rownames(results_df)[which.max(results_df$logLik)]

# Print the results data frame and best fitting distributions
results_df
cat("\nBest fitting distribution using AIC:", bestFitAIC, "\n")
cat("\nBest fitting distribution using BIC:", bestFitBIC, "\n")
cat("\nBest fitting distribution using Log-Likelihood:", bestFitLogLik, "\n")
## The best distribution here is weibull

# We can check graphically how the distribution is fitting to the survival data
require(flexsurv)
s <- with(survival_data_Gbimaculatus,Surv(Death_time,Censor))
swei  <- flexsurvreg(s ~ 1,dist="weibull",data=survival_data_Gbimaculatus)
sExp  <- flexsurvreg(s ~ 1,dist="exponential",data=survival_data_Gbimaculatus) 
plot(swei) +
lines(sExp, col="blue")
```

## Surv reg model

```{r }
model_fuma <- survreg(Surv(Death_time, Censor) ~ Treatment, data = survival_data_Fumagillin, dist="weibull")
summary(model_fuma)
Anova(model_fuma, type="II")
```
 
### Proper Modelling of Residual Variance

```{r }
RVAideMemoire::plotresid(model_fuma)
```

# Spores analysis

## Import data

```{r }
spores_data <- read_excel("Spores_data.xlsx")
colnames(spores_data)
```

## Convert variables to factor

```{r }
spores_data[c("Treatment", "Infection", "Density", "Sampling_Date", "Death_status", "Sex", "Mating_pair")] <- lapply(spores_data[c("Treatment", "Infection", "Density", "Sampling_Date", "Death_status", "Sex", "Mating_pair")], factor)

#Convert the "Spores" column to integers
spores_data$Estimated_Spores_insect <- floor(spores_data$Estimated_Spores_insect)
```

## Reorganise levels of Treatment

```{r }
spores_data$Treatment <- factor(spores_data$Treatment,levels = c("N_8","Y_8","N_16","Y_16","N_32","Y_32"))
```

## Data summary

```{r }
str(spores_data)
summary(spores_data)

# Calculate prevalence of infection by treatment
prevalence <- spores_data %>%
  group_by(Treatment) %>%
  summarise(Prevalence = sum(Estimated_Spores_insect > 0) / n())  # Number of infected samples (Spores > 0) divided by samples per treatment

# Calculate mean and standard error of spores per treatment
mean_se_spores <- spores_data %>%
  group_by(Treatment) %>%
  summarise(Mean_Spores = mean(Estimated_Spores_insect, na.rm = TRUE),  # Calculate mean spores, ignoring NA values
            SE_Spores = sd(Estimated_Spores_insect, na.rm = TRUE) / sqrt(n()), # Calculate standard error of spores (SD / sqrt(n)), ignoring NA values
            Count = n())  # Count the number of observations per treatment

# Print results
print(prevalence)
print(mean_se_spores)
```

## Box plot of Spores by Treatment

```{r }
# Calculate the log transformation of Spores
spores_data$sqrt_Spores <- sqrt(spores_data$Estimated_Spores_insect)

boxplot_spores <- ggplot(spores_data, aes(x = Treatment, y = sqrt_Spores, color = Treatment),
  legend.title = "Exposure_Density",
  legend.labs = c("N_8","Y_8","N_16","Y_16","N_32","Y_32")) +
  geom_boxplot() +
  geom_point() +
  scale_color_manual(values = c("#c4c4ff","#ffc4c4","#0000ff","#ff4e4e","#000027", "#760000")) +
  labs(title = "Number of spores by treatment",
       x = "Treatment",
       y = "Square root of spores number") +
  theme_bw(base_family = "Times New Roman") + 
  theme(axis.text = element_text(size = 12), 
        axis.title = element_text(size = 12), 
        legend.title = element_text(size = 10, face = "bold"),
        plot.title = element_text(size = 12, face = "bold", hjust = 0.5))
boxplot_spores
```

## Box Plot Facet by Sex

```{r }
boxplot_faceted_spores <- boxplot_spores + facet_wrap(~ Sex) +
  theme(strip.text.x = element_text(size=12),
  axis.text.x = element_text(angle = 50, vjust = 1, hjust = 1, size = 12)) +
  labs(colour = "Exposure_Density")
boxplot_faceted_spores
```

## Create histogram

```{r}
histogram_spores <- ggplot(spores_data, aes(x = sqrt_Spores
                                              #Treatment, y = sqrt_Spores
                                            #, fill = Treatment)
                           )) +
  geom_histogram() +
  #geom_histogram(stat="identity", binwidth = 0.1, color = "black") +
  #scale_fill_manual(values = c("#c4c4ff","#ffc4c4","#0000ff","#ff4e4e","#000027", "#760000")) +
  labs(title = "Number of spores",
       x = "Square root of spores number",
       y = "Frequencies") + 
  theme_bw(base_family = "Times New Roman") + 
  theme(axis.text = element_text(size = 12), 
        axis.title = element_text(size = 12),
        axis.text.x = element_text(angle = 50, vjust = 0.6), 
        plot.title = element_text(size = 12, face = "bold", hjust = 0.5))
histogram_spores
```

## Generalised Linear Model (glm) for quantitative discrete spore variable

```{r }
model_spores <- glm(Estimated_Spores_insect ~ Density, data = spores_data, family = quasipoisson(link = "log"))

model_spores1 <- glm(Estimated_Spores_insect ~ Density * Sex, data = spores_data, family = quasipoisson(link = "log"))

# Terms fit
anova_spores <- anova(model_spores, model_spores1)
## Model 1 has a lower deviance
#Extract p-value
1-pchisq( abs(anova_spores$Deviance[2]), abs(anova_spores$Df[2]))
## p-value = 0
```

### Check residuals and compare the model

```{r }
RVAideMemoire::plotresid(model_spores1)
summary(model_spores1)
Anova(model_spores1, type="II")
```

### Test significance of each term of the model with Anova

```{r }
pairwise_comp_sp <- emmeans(model_spores,pairwise ~ Density,type="link")
pairwise_comp_sp
cld(pairwise_comp_sp,adjust="fdr",Letters=letters)
```

# Fecundity analysis

## Import data

```{r }
fecundity_data <- read_excel("Fecundity_data.xlsx")
colnames(fecundity_data)
```

## Convert factors to appropriate levels

```{r }
fecundity_data[c("Treatment", "Infection", "Density", "Time_EggLaying")] <- lapply(fecundity_data[c("Treatment", "Infection", "Density", "Time_EggLaying")], factor)
```

## Reorganise levels of Treatment

```{r }
fecundity_data$Treatment <- factor(fecundity_data$Treatment,levels = c("N_8","Y_8","N_16","Y_16","N_32","Y_32"))
```

## Data summary

```{r }
str(fecundity_data)
summary(fecundity_data)

# Calculate mean and standard deviation of eggs per treatment
mean_sd_eggs <- fecundity_data %>%
  group_by(Treatment) %>%
  summarise(Mean_eggs = mean(Egg_Count, na.rm = TRUE),  
            SE_eggs = sd(Egg_Count, na.rm = TRUE) / sqrt(n()),
            Count = n())  # Count the number of observations per treatment

# Print results
print(mean_sd_eggs)
```

## Box plot of Spores by Treatment

```{r }
# Calculate the log transformation of Spores
#spores_data$sqrt_Spores <- sqrt(spores_data$Spores)

boxplot_eggs <- ggplot(fecundity_data, aes(x = Treatment, y = Egg_Count, color = Treatment),
  legend.title = "Exposure_Density",
  legend.labs = c("N_8","Y_8","N_16","Y_16","N_32","Y_32")) +
  geom_boxplot() +
  geom_point() +
  scale_color_manual(values = c("#c4c4ff","#ffc4c4","#0000ff","#ff4e4e","#000027", "#760000")) +
  labs(title = "Number of eggs by treatment",
       x = "Treatment",
       y = "Number of eggs") +
  theme_bw(base_family = "Times New Roman") + 
  theme(axis.text = element_text(size = 12), 
        axis.title = element_text(size = 12), 
        legend.title = element_text(size = 10, face = "bold"),
        plot.title = element_text(size = 12, face = "bold", hjust = 0.5))
boxplot_eggs
```

## Generalised Linear Model (glm) for quantitative discrete egg variable

```{r }
hist(fecundity_data$Egg_Count)
model_eggs <- glm(Egg_Count ~ Infection * Density, data = fecundity_data, family = quasipoisson(link = "log"))

model_eggs1 <- glm(Egg_Count ~ Infection * Density, data = fecundity_data, family = poisson(link = "log"))

anova(model_eggs, model_eggs1)
```

### Check residuals and compare the models and test significance of each term of the model with Anova

```{r }
RVAideMemoire::plotresid(model_eggs)
summary(model_eggs)
Anova(model_eggs, type="II")
```

### Pairwise comparison

```{r} 
pairwise_comp_egg <- emmeans(model_eggs,pairwise ~ Infection|Density,type="response")
pairwise_comp_egg
cld(pairwise_comp_egg,adjust="fdr",Letters=letters)
```

## Hatching success

```{r}
# First, convert Hatching_date to a logical variable indicating hatching success
fecundity_data$Hatching_success <- !is.na(fecundity_data$Hatching_date)

# Then calculate hatching success percentage for each treatment level
hatching_summary <- aggregate(Hatching_success ~ Treatment, data = fecundity_data, FUN = function(x) {
  mean(x) * 100
})

# Print the summary table
print(hatching_summary)

# Create a contingency table for hatching success by treatment
cont_table <- table(fecundity_data$Hatching_success, fecundity_data$Treatment)

# Perform chi-square test
chisq_result <- chisq.test(cont_table)
print(chisq_result)
```

# Pilot assay (June 2023)

## Import data

```{r }
survival_data_pilot <- read_excel("GbdensityJune23_survival_data.xlsx")
colnames(survival_data_pilot)
```

## Convert factors to appropriate levels

```{r }
survival_data_pilot[c("Treatment", "Infection", "Density")] <- lapply(survival_data_pilot[c("Treatment", "Infection", "Density")], factor)
```

## Reorganise levels of Treatment variable and convert variables to factors

```{r }
survival_data_pilot$Treatment <- factor(survival_data_pilot$Treatment,levels = c("N_10","Y_10","N_30","Y_30","N_70","Y_70"))
```

## Data summary

```{r }
str(survival_data_pilot)
summary(survival_data_pilot)
```

## Check mortality

```{r }
total_pilot <- survival_data_pilot %>% group_by(Treatment) %>% tally() #total number of accessions
mortality_pilot <- survival_data_pilot %>% group_by(Treatment, Censor) %>% tally() #total number of deaths 
mortality_pilot <-subset(mortality_pilot, Censor!=0)
colnames(total_pilot)[2] <- "total"
colnames(mortality_pilot)[3] <- "mortality"
pilot_check <- merge(total_pilot, mortality_pilot, by="Treatment", all = TRUE, fun=mean)
pilot_check$percent <- (pilot_check$mortality / pilot_check$total) *100 #calculate percentage of mortality
print(pilot_check)
```

## Surv curve Treatment

```{r }
surv_pilot <- survfit(Surv(Death_time, Censor) ~ Treatment, data = survival_data_pilot)
surv_plot_pilot <- ggsurvplot(
  fit = surv_pilot,
  title = "Cricket survival per treatment",
  surv.scale = "percent",
  legend.title = "Exposure_Density",
  legend.labs = c("N_10","Y_10","N_30","Y_30","N_70","Y_70"),
  palette = c("#c4c4ff","#ffc4c4","#0000ff","#ff4e4e","#000027","#760000"),
  linetype = c(1, 2, 1, 2, 1, 2),
  ggtheme = theme_bw(base_family = "Times New Roman") +
    theme(
      axis.text = element_text(size = 12), 
      axis.title = element_text(size = 12, vjust = 0.5),
      legend.text = element_text(size = 10),
      legend.title = element_text(size = 10, face = "bold"),
      legend.direction = "horizontal",
      plot.title = element_text(size = 12, face = "bold", hjust = 0.5)))
surv_plot_pilot
```

## LT50 table

```{r }
summary(surv_pilot)$table
```

## Determine if we have a constant risk or not

```{r }
RVAideMemoire::plotsurvivors(survival_data_pilot$Death_time, survival_data_pilot$Censor)
#The instantaneous risk is considered constant if the points on the survival curve are roughly aligned on a straight line. Here, the risk seems constant, but with a drop at the end.
```

## Surv reg model

### Find the best distribution for our survival data

```{r}
# Create vector of parametric distributions to test
distList <- c(distList <- c("extreme", "logistic", "gaussian", "weibull", "exponential", "rayleigh", "lognormal", "loglogistic", "t"))

# Function fits each distribution and extracts AIC, BIC, log-likelihood values
## The distribution is estimated on the event data without any covariate (~ 1)
fit_dist <- function(dist) {
  tmp <- survreg(Surv(Death_time, Censor) ~ 1, data = survival_data_pilot, dist = dist)
  c(AIC(tmp), BIC(tmp), as.numeric(logLik(tmp)))
}

# Apply above function to each distribution in the distList
results_list <- lapply(distList, fit_dist)

# Convert the above list of results to a data frame
results_df <- data.frame(t(matrix(unlist(results_list), nrow = 3)))
colnames(results_df) <- c("aic", "bic", "logLik")
rownames(results_df) <- distList

# Find the distribution with the lowest AIC, BIC, and logLik values
bestFitAIC <- rownames(results_df)[which.min(results_df$aic)]
bestFitBIC <- rownames(results_df)[which.min(results_df$bic)]
bestFitLogLik <- rownames(results_df)[which.max(results_df$logLik)]

# Print the results data frame and best fitting distributions
results_df
cat("\nBest fitting distribution using AIC:", bestFitAIC, "\n")
cat("\nBest fitting distribution using BIC:", bestFitBIC, "\n")
cat("\nBest fitting distribution using Log-Likelihood:", bestFitLogLik, "\n")
## The best distribution here is lognormal

# We can check graphically how the distribution is fitting to the survival data
require(flexsurv)
s <- with(survival_data_pilot,Surv(Death_time,Censor))
sLnorm  <- flexsurvreg(s ~ 1,dist="lnorm",data=survival_data_pilot)
sExp  <- flexsurvreg(s ~ 1,dist="exponential",data=survival_data_pilot) 
plot(sLnorm) +
lines(sExp, col="blue")
```

### Model

```{r }
model_pilot <- survreg(Surv(Death_time, Censor) ~ Infection * Density, data = survival_data_pilot, dist="lognormal")
summary(model_pilot)
```
 
### Proper Modelling of Residual Variance

```{r }
RVAideMemoire::plotresid(model_pilot)
```


### Test significance of each term of the model with Anova

```{r }
Anova(model_pilot, type="II")
```

### Pairwise comparison

```{r} 
pairwise_comp_p1 <- emmeans(model_pilot,pairwise ~ Infection,type="response")
pairwise_comp_p1
cld(pairwise_comp_p1,adjust="fdr",Letters=letters)

pairwise_comp_p2 <- emmeans(model_pilot,pairwise ~ Density,type="response")
pairwise_comp_p2
cld(pairwise_comp_p2,adjust="fdr",Letters=letters)

pairwise_comp_p3 <- emmeans(model_pilot,pairwise ~ Infection|Density,type="response")
pairwise_comp_p3
cld(pairwise_comp_p3,adjust="fdr",Letters=letters)
```


# Feed and feces analysis from Pilot assay

## Import data

```{r }
feed_feces_data_pilot <- read_excel("Feed_Feces_pilot_data.xlsx")
colnames(feed_feces_data_pilot)
```

## Reorganise levels of Treatment variable and convert variables to factors
## Here the experimental box was added as random factor, as multiple measurments (paired serie) were recorded on it

```{r }
feed_feces_data_pilot[c("Treatment", "Box_Id", "Infection", "Density", "Dpi")] <- lapply(feed_feces_data_pilot[c("Treatment", "Box_Id", "Infection", "Density", "Dpi")], factor)

feed_feces_data_pilot$Treatment <- factor(feed_feces_data_pilot$Treatment,
                                                 levels = c("N_10","Y_10","N_30","Y_30","N_70","Y_70"))
str(feed_feces_data_pilot)

# Subset the data to remove rows where Dpi = 2
feed_feces_data_pilot <- feed_feces_data_pilot[feed_feces_data_pilot$Dpi != 1, ]
# Remove unused levels of Dpi
feed_feces_data_pilot$Dpi <- droplevels(feed_feces_data_pilot$Dpi)
str(feed_feces_data_pilot)

```

## Calculate mean and standard deviation for each response variable by treatment

```{r }
summary_data <- aggregate(cbind(Feces_ind, Feed_ind) ~ Treatment , feed_feces_data_pilot, FUN = function(x) c(mean = mean(x), sd = sd(x)))
print(summary_data)
```

## Correlation between consumed feed and Faeces

```{r }
correlation_plot <- ggplot(feed_feces_data_pilot, aes(x = Feces_ind, y = Feed_ind)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Correlation between consumed Feed and Faeces",
       x = "Faeces weight (mg)",
       y = "Feed Consumed per individual (mg)") +
  theme(plot.title = element_text(face = "bold", size = 12, hjust = 0.5),
        axis.title.y = element_text(size = 12), 
        axis.text.x = element_text(angle = 50, vjust = 1, hjust = 1, size = 12),
        legend.position = "none")
correlation_plot
```

### Compute correlation coefficient

```{r }
correlation <- cor(feed_feces_data_pilot$Feed_ind, feed_feces_data_pilot$Feces_ind)
# Conduct Pearson's correlation test
cor_test <- cor.test(feed_feces_data_pilot$Feed_ind, feed_feces_data_pilot$Feces_ind)
# Print correlation coefficient
print(paste("Correlation coefficient:", correlation))
# Print correlation test results
print(cor_test)
```

## Histogram and boxplot of Feed Consumed by Treatment and Dpi

```{r }
# Filter out zero values
feed_feces_data_filtered <- feed_feces_data_pilot %>% filter(Feed_ind > 0)

# Create histogram with log-transformed Feed and by Dpi
histogram_feed_pilot <- ggplot(feed_feces_data_filtered, aes(x = Treatment, y = log(Feed_ind), fill = Treatment)) +
  geom_histogram(stat="identity", binwidth = 0.1, color = "black") +
  scale_fill_manual(values = c("#c4c4ff","#ffc4c4","#0000ff","#ff4e4e","#000027", "#760000")) +
  labs(title = "Feed consumed by treatment and Dpi",
       x = "Treatment",
       y = "Log Feed Consumed per individual (mg)") + 
  facet_wrap(~Dpi) +
  theme_bw(base_family = "Times New Roman") + 
  theme(axis.text = element_text(size = 12), 
        axis.title = element_text(size = 12),
        axis.text.x = element_text(angle = 50, vjust = 0.6), 
        plot.title = element_text(size = 12, face = "bold", hjust = 0.5))
histogram_feed_pilot

# Boxplot with all Dpi
boxplot_feed_pilot <- ggplot(feed_feces_data_pilot, aes(x = Treatment, y = Feed_ind, color = Treatment),
  legend.title = "Exposure_Density") +
  geom_boxplot() +
  geom_point() +
  scale_color_manual(values = c("#c4c4ff","#ffc4c4","#0000ff","#ff4e4e","#000027", "#760000")) +
  labs(title = "Feed consumed by treatment",
       x = "Treatment",
       y = "Feed Consumed per individual (mg)") +
  theme_bw(base_family = "Times New Roman") + 
  theme(axis.text = element_text(size = 12), 
        axis.title = element_text(size = 12), 
        plot.title = element_text(size = 12, face = "bold", hjust = 0.5))
boxplot_feed_pilot
```

## Feed model

```{r }
model_feed_pilot <- lmer(Feed_ind ~ Infection * Density + (1|Box_Id), data = feed_feces_data_pilot) #The box id is a random effect (different box measured across different time points)

model_feed_pilot_dpi <- lmer(Feed_ind ~ Infection * Density + Dpi + (1|Box_Id), data = feed_feces_data_pilot)
```

#### Check residuals

```{r }
RVAideMemoire::plotresid(model_feed_pilot)
summary(model_feed_pilot)

RVAideMemoire::plotresid(model_feed_pilot_dpi)
summary(model_feed_pilot_dpi)
```

#### Test significance of each term of the model with Anova

```{r }
Anova(model_feed_pilot, type="II")
Anova(model_feed_pilot_dpi, type="II")
```

# Emergence Analysis Pilot assay

## Import data for both sexes

```{r }
emergence_data_Gbimaculatus_pilot <- read_excel("AdultEmergence_pilot_data.xlsx")
colnames(emergence_data_Gbimaculatus_pilot)
```

### Convert factors to appropriate levels

```{r }
emergence_data_Gbimaculatus_pilot[c("Treatment","Infection","Density","Sex")] <- lapply(emergence_data_Gbimaculatus_pilot[c("Treatment","Infection","Density", "Sex")],factor)
```

### Reorganise levels of Treatment variable and convert variables to factors

```{r }
emergence_data_Gbimaculatus_pilot$Treatment <- factor(emergence_data_Gbimaculatus_pilot$Treatment,levels = c("N_10","Y_10","N_30","Y_30","N_70","Y_70"))
```

### Data summary

```{r }
str(emergence_data_Gbimaculatus_pilot)
```

## Create Male dataset

```{r }
emergence_male_Gbimaculatus_pilot <- filter(emergence_data_Gbimaculatus_pilot, Sex == "Male" | Sex == "NA")
emergence_male_Gbimaculatus_pilot$Sex <- droplevels(emergence_male_Gbimaculatus_pilot$Sex)
str(emergence_male_Gbimaculatus_pilot)
```

## Create Female dataset

```{r }
emergence_female_Gbimaculatus_pilot <- filter(emergence_data_Gbimaculatus_pilot, Sex == "Female" | Sex == "NA")
emergence_female_Gbimaculatus_pilot$Sex <- droplevels(emergence_female_Gbimaculatus_pilot$Sex)
str(emergence_female_Gbimaculatus_pilot)
```

## Calculate proportions and chi-square test

```{r echo=TRUE, message=FALSE}
emergence_table_pilot <- emergence_data_Gbimaculatus_pilot %>%
  group_by(Treatment,Infection,Density) %>%
  summarise(y1 = sum(Censor == 1), #Number of emerged crickets
            y2 = sum(Censor == 0)   #Number of non-emerged crickets
  ) %>%
  mutate(y1_prop = y1 / (y1 + y2), #Proportion of emerged crickets
         y2_prop = y2 / (y1 + y2))   #Proportion of non-emerged crickets

print(emergence_table_pilot)

# Create the contingency table
contingency_table <- as.matrix(emergence_table_pilot[, c("y1", "y2")])
rownames(contingency_table) <- paste(emergence_table_pilot$Treatment, emergence_table_pilot$Infection, sep = "_")
colnames(contingency_table) <- c("Emerged", "Non-emerged")

# Perform chi-square test
chisq_result <- chisq.test(contingency_table)
print

# Perform pairwise comparisons using fdr correction
pairwise_chisq <- pairwise.prop.test(x = contingency_table[, 1], n = rowSums(contingency_table), p.adjust.method = "fdr")
print(pairwise_chisq)
```

## LT50 table

```{r }
surv_emergence_pilot <- survfit(Surv(Emergence_time,Censor) ~ Treatment,data = emergence_data_Gbimaculatus_pilot)
summary(surv_emergence_pilot)$table
```


## Surv curve emergence

```{r }
surv_plot_emergence_pilot <- ggsurvplot(
  fit = surv_emergence_pilot,
  surv.scale = "percent",
  title = "Cricket emergence",
  ylab = "Cricket emergence",
  legend.title = "Exposure_Density",
  legend.labs = c("N_8","Y_8","N_16","Y_16","N_32","Y_32"),
  palette = c("#c4c4ff","#ffc4c4","#0000ff","#ff4e4e","#000027","#760000"),
  linetype = c(1, 2, 1, 2, 1, 2),
  xlim = c(14,34),
  break.x.by = 4,
  fun="event",
  #surv.median.line = "h",
  ggtheme = theme_bw(base_family = "Times New Roman") +
    theme(
      axis.text = element_text(size = 12), 
      axis.title = element_text(size = 12, vjust = 0.5),
      legend.text = element_text(size = 10),
      legend.title = element_text(size = 10, face = "bold"),
      plot.title = element_text(size = 12, face = "bold", hjust = 0.5)))

# Add line at 50% emergence
p2 <- surv_plot_emergence_pilot$plot +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "gold")
p2
```

## Surv curve male

```{r }
surv_male_pilot <- survfit(Surv(Emergence_time,Censor) ~ Treatment,data = emergence_male_Gbimaculatus_pilot)
surv_plot_male_pilot <- ggsurvplot(
  fit = surv_male_pilot,
  surv.scale = "percent",
  title = "Cricket male emergence",
  ylab = "Cricket emergence",
  legend.title = "Exposure_Density",
  legend.labs = c("N_10","Y_10","N_30","Y_30","N_70","Y_70"),
  palette = c("#c4c4ff","#ffc4c4","#0000ff","#ff4e4e","#000027","#760000"),
  linetype = c(1, 2, 1, 2, 1, 2),
  xlim = c(15,34),
  ylim = c(0,0.6),
  fun="event",
  ggtheme = theme_bw(base_family = "Times New Roman") +
    theme(
      axis.text = element_text(size = 12), 
      axis.title = element_text(size = 12, vjust = 0.5),
      legend.text = element_text(size = 10),
      legend.title = element_text(size = 10),
      plot.title = element_text(size = 12, face = "bold", hjust = 0.5)))
surv_plot_male_pilot
```

## Surv curve female

```{r }
surv_female_pilot <- survfit(Surv(Emergence_time,Censor) ~ Treatment,data = emergence_female_Gbimaculatus_pilot)
surv_plot_female_pilot <- ggsurvplot(
  fit = surv_female_pilot,
  surv.scale = "percent",
  title = "Cricket female emergence",
  ylab = "Cricket emergence",
  legend.title = "Exposure_Density",
  legend.labs = c("N_10","Y_10","N_30","Y_30","N_70","Y_70"),
  palette = c("#c4c4ff","#ffc4c4","#0000ff","#ff4e4e","#000027","#760000"),
  linetype = c(1, 2, 1, 2, 1, 2),
  xlim = c(15,34),
  ylim = c(0,0.6),
  fun="event",
  ggtheme = theme_bw(base_family = "Times New Roman") +
    theme(
      axis.text = element_text(size = 12), 
      axis.title = element_text(size = 12, vjust = 0.5),
      legend.text = element_text(size = 10),
      legend.title = element_text(size = 10),
      plot.title = element_text(size = 12, face = "bold", hjust = 0.5)))
surv_plot_female_pilot
```

## Arrange the surv plots together

```{r }
splots_pilot <- list(surv_plot_female_pilot, surv_plot_male_pilot)
arrange_pilot <- arrange_ggsurvplots(splots_pilot, print = TRUE, ncol = 2, nrow = 1)
arrange_pilot
```

## Generalized linear model – binomial family

```{r }
#Filter out NA values from the initial data set
#emergence_data_pilot_filtered <- emergence_data_Gbimaculatus_pilot %>% filter(Sex != "NA")

# Make a copy of your dataset
##emergence_data_copy <- emergence_data_Gbimaculatus_pilot
# Convert 'Sex' column to character type
##emergence_data_copy$Sex <- ##as.character(emergence_data_copy$Sex)
# Replace 'NA' with 'Unknown' in the 'Sex' column of the copied dataset
##emergence_data_copy$Sex[emergence_data_copy$Sex == "NA"] <- "Unknown"
# Convert 'Sex' column back to factor
##emergence_data_copy$Sex <- factor(emergence_data_copy$Sex)
##str(emergence_data_copy)

glm_model_pilot <- glm(Emergence_time ~ Infection * Density, data = (emergence_data_Gbimaculatus_pilot), family = Gamma(link = "inverse"))

summary(glm_model_pilot)

RVAideMemoire::plotresid(glm_model_pilot)
```

### Test significance of each term of the model with Anova and pairwise comparisons

```{r}
Anova(glm_model_pilot,type="II")

pairwise_comp_glmp <- emmeans(glm_model_pilot,pairwise ~ Infection,type="response")
pairwise_comp_glmp
cld(pairwise_comp_glmp,adjust="fdr",Letters=letters)

pairwise_comp_glmd <- emmeans(glm_model_pilot,pairwise ~ Density,type="response")
cld(pairwise_comp_glmd,adjust="fdr",Letters=letters)

pairwise_comp_glmp1 <- emmeans(glm_model_pilot,pairwise ~ Infection|Density,type="response")
cld(pairwise_comp_glmp1,adjust="fdr",Letters=letters)
```

### Plot emmeans

```{r}
# Create a sequence of values
#full_sequence <- seq(0, 1550, by = 50)

# Select every third value from the sequence
#selected_values <- full_sequence[seq(1, length(full_sequence), by = 3)]

#Plot emmeans data
pairwiseglm_plot <- plot(pairwise_comp_glmp1) +
  labs(title = "Pairwise Comparison of Estimated Means of Emergence Time",
       x = "Estimated Means of Emergence Time",
       y = "Infection") +
  #scale_x_continuous(breaks = selected_values) +
  theme_bw(base_family = "Times New Roman") +
  theme(
    axis.text = element_text(size = 12), 
    axis.title = element_text(size = 12, vjust = 0.5),
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 12),
    plot.title = element_text(size = 12, face = "bold", hjust = 0.5))

# Print the plot
print(pairwiseglm_plot)
```

## Event data model

```{r}
RVAideMemoire:: plotsurvivors(emergence_data_Gbimaculatus_pilot$Emergence_time, emergence_data_Gbimaculatus_pilot$Censor)
#Risk seems constant
```

### Find the best distribution for our survival data

```{r}
# Create vector of parametric distributions to test
distList <- c(distList <- c("extreme", "logistic", "gaussian", "weibull", "exponential", "rayleigh", "lognormal", "loglogistic", "t"))

# Function fits each distribution and extracts AIC, BIC, log-likelihood values
## The distribution is estimated on the event data without any covariate (~ 1)
fit_dist <- function(dist) {
  tmp <- survreg(Surv(Emergence_time, Censor) ~ 1, data = emergence_data_Gbimaculatus_pilot, dist = dist)
  c(AIC(tmp), BIC(tmp), as.numeric(logLik(tmp)))
}

# Apply above function to each distribution in the distList
results_list <- lapply(distList, fit_dist)

# Convert the above list of results to a data frame
results_df <- data.frame(t(matrix(unlist(results_list), nrow = 3)))
colnames(results_df) <- c("aic", "bic", "logLik")
rownames(results_df) <- distList

# Find the distribution with the lowest AIC, BIC, and logLik values
bestFitAIC <- rownames(results_df)[which.min(results_df$aic)]
bestFitBIC <- rownames(results_df)[which.min(results_df$bic)]
bestFitLogLik <- rownames(results_df)[which.max(results_df$logLik)]

# Print the results data frame and best fitting distributions
results_df
cat("\nBest fitting distribution using AIC:", bestFitAIC, "\n")
cat("\nBest fitting distribution using BIC:", bestFitBIC, "\n")
cat("\nBest fitting distribution using Log-Likelihood:", bestFitLogLik, "\n")
## The best distribution here is lognormal

# We can check graphically how the distribution is fitting to the survival data
require(flexsurv)
s <- with(emergence_data_Gbimaculatus_pilot,Surv(Emergence_time,Censor))
sLnorm  <- flexsurvreg(s ~ 1,dist="lnorm",data=emergence_data_Gbimaculatus_pilot)
sExp  <- flexsurvreg(s ~ 1,dist="exponential",data=emergence_data_Gbimaculatus_pilot) 
plot(sLnorm) +
lines(sExp, col="blue")
```

```{r}
model_surv_pilot <- survreg(Surv(Emergence_time, Censor) ~ Infection * Density, data = emergence_data_Gbimaculatus_pilot, dist="lognormal")

RVAideMemoire::plotresid(model_surv_pilot)

Anova(model_surv_pilot, type="II")

pairwise_comp_survp <- emmeans(model_surv_pilot,pairwise ~ Infection,type="response")
cld(pairwise_comp_survp,adjust="fdr",Letters=letters)

pairwise_comp_survp1 <- emmeans(model_surv_pilot,pairwise ~ Infection|Density,type="response")
cld(pairwise_comp_survp1,adjust="fdr",Letters=letters)
```

### Plot emmeans

```{r}
# Create a sequence of values
full_sequence <- seq(0, 1550, by = 50)

# Select every third value from the sequence
selected_values <- full_sequence[seq(1, length(full_sequence), by = 3)]

#Plot emmeans data
pairwise_plot <- plot(pairwise_comp_survp1) +
  labs(title = "Pairwise Comparison of Estimated Means of Emergence Time",
       x = "Estimated Means of Emergence Time",
       y = "Infection") +
  scale_x_continuous(breaks = selected_values) +
  theme_bw(base_family = "Times New Roman") +
  theme(
    axis.text = element_text(size = 12), 
    axis.title = element_text(size = 12, vjust = 0.5),
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 12),
    plot.title = element_text(size = 12, face = "bold", hjust = 0.5))

# Print the plot
print(pairwise_plot)
```

# Create a table that compiles all the above statistical outputs from the Anova objects

* Function to extract results

```{r}
get.univariate.results <- function(model) {
  anova_results <- car::Anova(model, type = "II") %>% broom::tidy()
  
  if ("lme" %in% class(model) || "lmerMod" %in% class(model) || "glmerMod" %in% class(model)) {
    random_effects <- paste(names(VarCorr(model)), collapse = ", ")
    variances <- paste(format(VarCorr(model), digits = 2), collapse = ", ")
  } else {
    random_effects <- NA
    variances <- NA
  }
  
  cbind(
    `Response (y)` = deparse(summary(model)$call$formula[[2]]),
    `Model` = as.character(summary(model)$call[[1]]),
    Formula = paste ("y ~", deparse(summary(model)$call$formula[[3]])),
    anova_results %>%
      dplyr::mutate(
        statistic = round(statistic, digits = 2),
        signif = gtools::stars.pval(p.value),
        p.value = round(p.value, digits = 3),
        p.value = format.pval(p.value, digits = 3)
      ),
    random_effects = random_effects,
    variances = variances
  ) %>% as.data.frame
}
```

* Create a list that contains all models

```{r}
sum.model <- list(glm_model1, model_weight1, model_feces_age, model_spores1, model_eggs)
```

* Apply the function to the list

```{r message=FALSE, warning=FALSE}
univariate.output <- purrr::map(.x = sum.model, .f = get.univariate.results) %>%
  dplyr::bind_rows()
univariate.output <- dplyr::rename(univariate.output, c(Factor = term, `Test value` = statistic, `P value`= p.value))

knitr::kable(univariate.output)
write.table(univariate.output, file = "models_stats.txt", sep = ";", dec = ".")
```